% ACKNOWLEDGEMENT PAGE
\thispagestyle{empty}
\newpage
\cleardoublepage\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgements}\index{Acknowledgements}
\begin{center}
{\Large \bf Acknowledgements}\\
\end{center}
\vspace{10pt}


\hspace{10mm}We are deeply indebted to our Internal Guide, Prof. Aparajita Biswal, for her constant guidance, motivation, suggestions, and encouragement throughout our project work, and for being instrumental in making our system more effective. She has devoted a significant amount of her valuable time to plan and discuss the project with us. Without her experience and insights, it would have been very difficult to accomplish this distinctive work. 

\par We are thankful to our Major Project Coordinators, Ms. Namrata Patel, Mr. Subham Upadhyay, of the Computer Science Engineering Department, Parul Institute of Engineering and Technology, Parul University, for directing us to carry out this project within the given timeframe and for guiding us in reading, searching, and writing a literature review. Their guidance regarding project aspects, generous assistance, and cooperation have been invaluable.

\par We would like to express our sincere gratitude to Dr. Shailendra Kumar Mishra, Head of Department, CSE, PIET, Parul University, for supervising us in this project work and providing us with the necessary support to bring our vision to fruition.

\par Last but not least, we would like to thank our college, the faculties of the CSE department, and our friends for helping us out in difficult situations during our project work. We are also grateful to our parents for supporting us throughout the completion of the hardware development and literature review.

\par Above all, we would like to thank the Almighty for giving us the strength to build the software and complete the project work with success.

\vspace{1.5cm}

\begin{flushright}
    \textbf{Santoki Soumil Shaileshbhai \\ Patel Krish Minolkumar \\ Raxitkumar Liladharbhai Gamdha \\Saish Kiran Waghmare}
\end{flushright}

\vspace{1pt}

\begin{flushright}
    \textbf{CSE, PIET \\ Parul University, \\ Vadodara} 
\end{flushright}
